---
title: "Analyze abundance data of Garrulus glandarius"
output: 
  html_document
editor_options: 
  chunk_output_type: console
always_allow_html: true
---
```{r, set-global-options}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```


```{r, echo=FALSE}
library("tidyverse")
library("here")
library("car")
library("report")
library("plotrix")
library("broom")
library("kableExtra")
library("ggstatsplot")
library("statsExpressions")
library("plotrix") #std.error 
library("flextable")
library("DT")
library("gtsummary")
```

## Prepare data
### Data sources

*Obsnev bird censuses* provided by [*OBSNEV*](https://obsnev.es/), realized in several transects distributed along Sierra Nevada. Temporal range from 2008 to 2020. The data were downloaded from new information system of OBSNEV (*i.e.* PostgreSQL [db01.obsnev.es](db01.obsnev.es))
  
### Prepare OBSNEV bird censuses 

- Raw data were downloaded from OBSNEV information system. The downloaded tables were: 
`contactos_paseriformes.csv`; `dicc_especies.csv`; `geo.csv`; `visitas.csv`. 

- The protocol sampling number for passerine in the database is 5. 

- Select the two locations: "Cortijo del Hornillo" (SANJUAN), "Robledal de Canar" (CANAR)

- Select only the taxonomic level = species (level > 6) 

- Filter out data contacted over 30 m 

```{r read-obsnev-tables}
contactos <- read_csv(here::here("data/birds/contactos_paseriformes.csv")) %>% 
  dplyr::select(-fcreacion, -fmodificacion)

dicc_sp <- read_csv(here::here("data/birds/dicc_especies.csv")) %>% 
  dplyr::select(idesp, nombre_cientifico, nivel)

# El protocolo de paseriformes es el 5
dicc_visita <- read_csv(here::here("data/birds/visitas.csv"), 
                        col_types = 
                          cols(.default ="?", 
                               idgeo = col_character(), 
                               fvisita = col_datetime(format="%Y-%m-%d %H:%M:%S"))) %>%
  filter(protocolo == 5) 
  
dicc_geo <- read_csv(here::here("data/birds/geo.csv"), 
                     col_types = cols(.default ="?", 
                               longitud_m = col_double()))

visita_geo <- 
  dicc_visita %>% inner_join(dicc_geo, by = "idgeo") %>% 
  dplyr::select(idvisitas, fvisita, nombre, longitud_m) 

dfraw <- contactos %>% 
  inner_join(visita_geo, by = "idvisitas") %>% 
  inner_join(dicc_sp, by = "idesp") %>% 
  mutate(year = lubridate::year(fvisita), 
         mes = lubridate::month(fvisita))

df <- dfraw %>% 
  filter(nombre %in% c("Cortijo del Hornillo", "Robledal de C치침ar")) %>% 
  mutate(habitat = case_when(
    stringr::str_detect(nombre, "Hornillo") ~ "SANJUAN",
    stringr::str_detect(nombre, "C치침ar") ~ "CANAR")) %>% 
  mutate(year = lubridate::year(fvisita), 
         mes = lubridate::month(fvisita),
         fecha = strftime(fvisita, format="%Y-%m-%d")) %>% 
  filter(nivel > 6) %>% 
  filter(desplazamiento < 31)

```


- Aggregate data to get the sum of all contacts for a specie recorded in the same visit of the same transect

- Select data from 2008 to 2014 

```{r, compute-abundance-obsnev}
garrulus <- df %>% 
  filter(nombre_cientifico == "Garrulus glandarius") %>% 
  group_by(fecha, year, mes, longitud_m, nombre, habitat, nivel) %>%
  summarise(total_ind = sum(numero)) %>%  
  mutate(den = round((total_ind * 10000 * 10 / (longitud_m * 60)),2)) %>% 
  ungroup() %>% 
  rename(localidad = habitat) %>% 
  dplyr::select(-nivel, -longitud_m, -nombre, -total_ind) %>% 
  filter(year > 2007) %>% 
  filter(year < 2015)
```

```{r}
report_text(garrulus)
```


### Aggregate data by year 
- Bird abundance (ind / 10 ha)

```{r}
garrulus_by_year <- garrulus %>% 
  group_by(localidad, year) %>% 
  summarise(mean = mean(den, na.rm = TRUE), 
            se = std.error(den), 
            n = length(den))

garrulus_by_year %>% 
  datatable(filter = "top") %>% 
  formatRound(columns=c('mean', 'se'), digits=2)
```

## Exploratory Plot 

```{r coloniza-garrulus, dev=c('png', 'pdf', 'svg', 'tiff')}
pos <- position_dodge(0.15)
plot_garrulus <- garrulus_by_year %>% 
  rename(site = localidad) %>% 
  ggplot(aes(x=year, y=mean, colour=site)) + 
  geom_point(position = pos) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                width=.1,position = pos) +
  geom_line(position = pos) + 
  theme_bw() +
  ylab("Eurasian jay abundance (birds/10ha)") + 
  xlab("Year") + 
  scale_color_manual(values=c("blue", "black")) + 
  theme(panel.grid = element_blank(),
        legend.position = c(.25,.85), 
        legend.title = element_blank()) 
print(plot_garrulus)
```

## Analysis of variance  
- The data are unbalanced (we used type III)

```{r}
garrulus$year <- as.factor(garrulus$year)
aov.unb <- aov(den ~ year * localidad, data = garrulus)
a <- car::Anova(aov.unb, type ="III", test="F")
report(a)
```

```{r}
knitr::kable(a, digits = c(3,0,3,4)) %>% 
  kable_styling()
```

```{r}
aa <- anova(aov.unb, test="F")
knitr::kable(aa, digits = c(0,3,3,3,4)) %>% kable_styling()
```

### Plot exploring bewteen sites 

```{r}
ggbetweenstats(data=garrulus, 
               x=localidad, 
               y=den)
```

- Get mean values by locality 

```{r}
garrulus_by_year %>% 
  ungroup() %>% 
  rename(avg=mean) %>% 
  group_by(localidad) %>% 
    summarise(mean = mean(avg, na.rm = TRUE), 
            se = std.error(avg), 
            n = length(avg)) %>% 
  kable(digits = c(0,2,2,0)) %>% kable_styling()
```

