---
title: "Model Juvenile Abundance"
output: 
  html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
library("tidyverse")
library("here")
library("knitr")
library("kableExtra")
library("plotrix") #std.error 
library("rcompanion")
library("statmod")
library("tweedie")
# load functions from Zuur et al (2009) 
source(here::here("analysis/HighstatLibV6.R"))

# https://stackoverflow.com/questions/49836521/trouble-installing-glmmadmb-in-r
library('glmmADMB') # See Zuur et al. 2013. p146 
library('lattice')
library('MASS') #for negative binomial
library("performance")
library("DHARMa")
library("MuMIn")
library("broom")
```

## Prepare Data 

```{r, readData}
data <- read_csv(here::here("data/tree_data_crops.csv")) %>%
  mutate(sp = case_when(
    especie == "roble" ~ "qp",
    especie == "encina" ~ "qi",
    especie == "pino" ~ "pi"
  ))

# Add data of the transects selected
transectos <- read_csv(here::here("data/transectos_crops.csv"))

info_transectos <- transectos %>%
  dplyr::select(nombre, tipo, localidad, crop_code, altitud)

juvenile_abundance <- data %>%
  group_by(sp, nombre) %>%
  count() %>%
  pivot_wider(names_from = sp, values_from = n, names_prefix = "juv_") %>%
  rowwise() %>%
  mutate(
    ntotal = sum(juv_qp, juv_qi, juv_pi, na.rm = TRUE),
    nquercus = sum(juv_qp, juv_qi, na.rm = TRUE),
    juvenile_ab = round(ntotal * 100 / 300, 2),
    juvenileQuer_ab = round(nquercus * 100 / 300, 2),
    juvenileQp_ab = round(juv_qp * 100 / 300, 2)
  )

df <- info_transectos %>%
  left_join(juvenile_abundance) %>%
  mutate_all(~ replace_na(., 0))


qp <- df %>%
  dplyr::select(nombre, tipo, localidad, altitud, crop_code, ab = juvenileQp_ab) %>%
  mutate(
    nombre = as.factor(nombre),
    tipo = as.factor(tipo),
    localidad = as.factor(localidad),
    crop_code = as.factor(crop_code)
  )

```

# Modelization 

- Negative binomial, Poisson or Tweedie. 

- First we modelized a Negative Binomial, because Poisson model is overdispersed 

```{r modelNB}
nb <-MASS::glm.nb(ab ~ tipo*localidad, data = qp, link = "log")
```

- Check distribution of the data, using `performance::check_distribution`

```{r}
performance::check_distribution(nb)
```


```{r}
plot(performance::check_distribution(nb))
```

- We will model with Tweedie distribution, with $y \sim Tw_{\xi}(\mu, \phi)$

## Tweedie GLM
- First we estimate $\xi$ using `tweedie::tweedie.profile` function.

```{r}
set.seed(1234)
# xi parameter 
out <- tweedie.profile(ab ~ tipo*altitud, data = qp, 
                       do.plot = TRUE,
                       xi.vec = seq(1.05, 1.8, by= 0.05))
```

```{r}
t(t(c("MLE of xi" = out$xi.max, "lower CI 95%" = out$ci[1], "upper CI 95 %" = out$ci[2]))) %>% as.data.frame() %>% rename("Maximum Likelihood Estimation" = V1)
```

```{r}
xi.est <- round(out$xi.max, 2)
```

- Now, we modeled the Tweedie GLM using the estimate of the index parameter ($\xi = $ `r xi.est`) 

```{r}
m <- glm(ab ~ tipo*localidad, data = qp, family = tweedie(var.power = xi.est, link.power = 0))
```

- Check model diagnostics 

```{r}
performance::check_model(m)
```

### Model selection  
```{r}
library(performance)
mfull <- glm(ab ~ tipo*localidad, data = qp, family = tweedie(var.power = xi.est, link.power = 0))
mtl <- glm(ab ~ tipo + localidad, data = qp, family = tweedie(var.power = xi.est, link.power = 0))
mt <- glm(ab ~ tipo, data = qp, family = tweedie(var.power = xi.est, link.power = 0))
ml <- glm(ab ~ localidad, data = qp, family = tweedie(var.power = xi.est, link.power = 0))
mnull <- glm(ab ~ 1, data = qp, family = tweedie(var.power = xi.est, link.power = 0))

r2 <- function(mimodelo){round(((mimodelo$null.deviance - mimodelo$deviance) / mimodelo$null.deviance),3)}

models_selection <- 
model.sel(mfull, mtl, mt, ml, mnull, rank = AICtweedie) %>% as.data.frame() %>% 
    mutate(model.name = c(
      "~ habitat type + site + habitat type:site", 
      "~ habitat type + site", 
      "~ habitat type", 
      "~ site",
      "null model"), 
           f = c(mfull$formula, mtl$formula, mt$formula, ml$formula, mnull$formula), 
           r2nagelkerke = round(
             c(r2_nagelkerke(mfull), r2_nagelkerke(mtl), r2_nagelkerke(mt), r2_nagelkerke(ml),r2_nagelkerke(mnull)),3),
           r2 = round(
             c(r2(mfull), r2(mtl), r2(mt), r2(ml), r2(mnull)),3),
           logLiktweedie = c(logLiktweedie(mfull),
                             logLiktweedie(mtl),
                             logLiktweedie(mt),
                             logLiktweedie(ml),
                             logLiktweedie(mnull)
                             )) %>% 
  dplyr::select(-f, -localidad, -tipo, -`localidad:tipo`, -family,-`(Intercept)`, -logLik, -weight) %>% 
  relocate(model.name, df, logLiktweedie) 

write.table(models_selection, 
            here::here("ms/table_model_selection.csv"))

```


```{r}
options(knitr.kable.NA = '') # suppress showing NA values in table
kable(models_selection, format = "html",  
      digits = c(0, 0, 2, 2, 2, 3,3), align = "rrrrrrr", row.names = FALSE)  %>%
  kable_styling(position = "left")
```

### Anova table

```{r}
library(car)
library(kableExtra)
a <- Anova(mfull, test="F") %>% as.data.frame()
names(a) <- c("SS", "df", "F", "p-value")

write.table(a, here::here("ms/table_anova.csv"))
```

```{r}
options(knitr.kable.NA = '') # suppress showing NA values in table
kable(a, format = "html",  
      digits = c(3, 0, 2, 4), align = "rrrr")  %>%
  kable_styling(position = "left")
```


- Report
```{r, warning=FALSE}
library("report")
report(mfull)
report_performance(mfull)
report_statistics(mfull)
```


### Plot 
```{r}
library(ggpubr)
library(broom)
library(plotrix)
library(rcompanion)
library(multcompView)
library(multcomp)
library(ggrepel)

ab_mean <- qp %>% 
  group_by(localidad, tipo) %>% 
  summarise(
    mean = mean(ab, na.rm = TRUE),
    sd = sd(ab, na.rm = TRUE),
    se = plotrix::std.error(ab, na.rm = TRUE)
  ) 

ab_boot <- groupwiseMean(ab ~ localidad + tipo,
              data = qp,
              conf = 0.95,
              digits = 3,
              R   = 10000,
              boot        = TRUE,
              traditional = FALSE,
              normal      = FALSE,
              basic       = FALSE,
              percentile  = FALSE,
              bca         = TRUE) %>% 
  mutate(site = 
           case_when(localidad == "CANAR" ~ "CA", 
                     TRUE ~ "SJ"),
         type = 
           case_when(tipo == "CLARO" ~ "Cropland", 
                     tipo == "CLARO_BORDE" ~ "Edge",
                     TRUE ~ "Forest")) %>% 
  inner_join((qp %>% group_by(localidad, tipo) %>% 
  summarise(se = round(plotrix::std.error(ab, na.rm = TRUE),2)))) %>% 
  relocate(localidad:Mean, se) %>% ungroup() %>% as.data.frame()

## Add letters 

### Compare bewteen sites by type
library(lsmeans)

letter_type <- 
  cld(lsmeans(mfull, ~ tipo|localidad),
    level = .95,
    Letters=letters,      ### Use lower-case letters for .group
    adjust="tukey",        ### Tukey-adjusted comparisons 
) %>% data.frame()

ab_boot <- ab_boot %>% inner_join(letter_type) %>% 
  rename(letra = `.group`)

l_site <- ab_boot %>% group_by(type) %>% 
  summarise(y = mean(Mean)) %>% 
  mutate(dif = 
           case_when(type == "Cropland" ~ "***",
                     type == "Edge" ~ "***",
                     type == "Forest" ~ ""))


pd <- position_dodge(.4)
ab_interaction <- 
  ggplot(data=ab_boot, 
         aes(x=site, y=Mean, shape=type, group=type)) +
    geom_point(color='black', size=4, position=pd) +
    geom_errorbar(aes(ymin=Bca.lower, ymax=Bca.upper), 
                  color='black', width=.2, position=pd) +
    geom_line(position=pd) +
    xlab("site") + 
  ylab(expression(paste("Juvenile abundance (ind/100 ",m^2,")"))) + 
    theme_bw() +
    guides(shape=guide_legend(title="")) + 
  theme(panel.grid = element_blank()) + 
  geom_text_repel(nudge_x = 0.2,
                  nudge_y = 0.1,
                  direction = "both",
                  aes(label=letra), 
                  segment.color = NA) +
  geom_text(data=l_site, aes(y=y, x=1.6, label=dif))+
  theme(axis.title=element_text(size = rel(1.7)),
          axis.text=element_text(size = rel(1.7)),
          legend.title=element_text(size = rel(1.3)),
          legend.text=element_text(size = rel(1.3)))
```

```{r}
write.table(ab_boot, here::here("ms/table_means.csv"))
```

```{r, ab-interaction, dev=c('png', 'pdf', 'svg', 'tiff')}
print(ab_interaction)
```

```{r}
ggsave(ab_interaction, filename = here::here("ms/juvenile_interaction.pdf"))
```











- post hocs 
```{r}
pairs(emmeans(mfull, c("tipo"), by="localidad")) %>% 
  data.frame() %>% 
  kable(format = "html",  
      digits = c(0,0,3, 3, 0, 3, 4), align = "rrrrrrr")  %>%
  kable_styling(position = "left")
```

```{r}
pairs(emmeans(mfull, c("localidad"), by="tipo")) %>% data.frame() %>% 
    kable(format = "html",  
      digits = c(0,0,3, 3, 0, 3, 4), align = "rrrrrrr")  %>%
  kable_styling(position = "left")

```

```{r, echo = FALSE, eval=FALSE}
data.frame(multcomp::cld(
  pairs(emmeans(mfull, "localidad", by="tipo"), 
    Letters=letters,
    alpha = 0.05))) %>% 
  rename(comp_type = `.group`)
```





   








