---
title: "Compare Forest Structure"
output: 
  html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Intro

- The aim of this code is to compare the forest structure of the two study sites. 

- We selected several dependent variables and performed non-parametric analysis comparing the value of the variables between sites 

- Dependent variables: 
  
  - Total abundance 
  - Composition (% of *Q. pyrenaica*)
  - Juvenile abundance
  - Tree-height statistics (mean, median, maximum)
  - Basal Area (m^2^ ha^-1^)


```{r, pkg}
library(tidyverse)
library(here)
library(knitr)
library(kableExtra)
library(ggstatsplot)
library(statsExpressions)
library(plotrix) #std.error 
library("flextable")
```

## Prepare data 
```{r, prepareData}
# Select data from forest
# Remove data without values in altura field
forest <- read_csv(here::here("data/tree_data_crops.csv")) %>% 
  filter(tipo == 'ROBLEDAL') %>% 
  filter(!is.na(altura))  

# composition and abundance (ind /ha)
composition <- forest %>% 
  group_by(crop_code, localidad, especie, nombre) %>% 
  count() %>% 
  pivot_wider(names_from = especie, values_from = n) %>% 
  rowwise() %>% 
  mutate(ntotal = sum(encina,roble, na.rm = TRUE),
         per_qp = round(roble/ntotal*100, 2), 
         per_qi = ifelse(
           is.na(encina), 0, 
           round(encina/ntotal*100, 2)), 
         abundance = round(ntotal * 10000/300, 2)) 

# abundance Qp (ind/ha)
ab <- forest %>% 
  group_by(crop_code, localidad, especie, nombre, tree_type) %>% 
  count() %>% 
  mutate(n = n*10000/300) %>% 
  pivot_wider(names_from = tree_type, values_from = n) 

ab_roble <- ab %>% filter(especie=="roble") %>% 
  mutate(ab_juvenile = replace_na(juvenile, 0),
         ab_adult = replace_na(adult, 0)) %>% 
  dplyr::select(-juvenile, -adult,  -especie)

# height
tree_height <- forest %>% 
  filter(especie == "roble") %>% 
  group_by(crop_code, localidad, nombre) %>% 
  summarise(tree_height_max = max(altura, na.rm = TRUE)/100,
            tree_height_mean = mean(altura, na.rm = TRUE)/100,
            tree_height_median = median(altura, na.rm = TRUE)/100,
            tree_height_q75 = quantile(altura, 0.75)/100,
            tree_height_q90 = quantile(altura, 0.9)/100)


# BASAL AREA (m2/ha)
ba <- forest %>% 
  filter(especie == "roble") %>% 
  # filter(tree_type == "adult") %>% 
  mutate(ba_tree = (pi*diametro_base*diametro_base)/40000) %>% 
  group_by(crop_code, localidad, nombre) %>% 
  summarise(ba = ((sum(ba_tree))*10000)/300)

forest_str <- composition %>% 
  dplyr::select(crop_code, localidad, nombre, per_qp, abundance) %>% 
  inner_join(ab_roble) %>% 
  inner_join(tree_height) %>% 
  inner_join(ba) %>% 
  dplyr::select(-especie)

write_csv(forest_str, here::here("data/forest_structure.csv"))
```


## Get summary by locality
```{r}
ml <- forest_str %>% 
  group_by(localidad) %>% 
  summarise_at(vars(per_qp:ba), .funs = mean) %>% 
  pivot_longer(per_qp:ba, names_to = "variable", values_to = "mean") 

sl <- forest_str %>% 
  group_by(localidad) %>% 
  summarise_at(vars(per_qp:ba), .funs = std.error) %>% 
  pivot_longer(per_qp:ba, names_to = "variable", values_to = "se") 

statby_localidad <- ml %>% dplyr::inner_join(sl) %>% 
  mutate(across(where(is.numeric), round, 2)) %>% 
  unite("value", mean:se, sep=" ± ")

# Apply test 
variables_interes <- forest_str %>% ungroup %>% dplyr::select(per_qp:ba) %>% names()

test_localidad <- c() 

for (i in 1:length(variables_interes)){
  test <- two_sample_test(data = forest_str, 
                  x = localidad, 
                  y = !!variables_interes[i], paired = FALSE, type = "np")
  
  test_localidad <- rbind(test_localidad, test)
}

test_localidad <- test_localidad %>% rename(variable = parameter1)


forest_str_by_localidad <- statby_localidad %>% 
  pivot_wider(values_from = value, names_from = localidad) %>% 
  inner_join(test_localidad) %>% 
  dplyr::select(-expression) 

forest_str_by_localidad %>% 
  flextable(col_keys = c("variable","CANAR", "SANJUAN", "statistic", "p.value")) %>% 
  colformat_double(digits = 3)

forest_str_by_localidad %>% write_csv(here::here("data/forest_str_by_localidad.csv"))

```


# Get summary by cropland
```{r}
m <- forest_str %>% 
  group_by(crop_code) %>% 
  summarise_at(vars(per_qp:ba), .funs = mean) %>% 
  pivot_longer(per_qp:ba, names_to = "variable", values_to = "mean") 

s <- forest_str %>% 
  group_by(crop_code) %>% 
  summarise_at(vars(per_qp:ba), .funs = std.error) %>% 
  pivot_longer(per_qp:ba, names_to = "variable", values_to = "se") 

statby_crop <- m %>% dplyr::inner_join(s) %>% 
  mutate(across(where(is.numeric), round, 2)) %>% 
  unite("value", mean:se, sep=" ± ")


# Apply test 
variables_interes <- forest_str %>% ungroup %>% dplyr::select(per_qp:ba) %>% names()

test_crop <- c() 

for (i in 1:length(variables_interes)){
  test <- oneway_anova(data = forest_str, 
                  x = crop_code, 
                  y = !!variables_interes[i], paired = FALSE, type = "np")
  
  test_crop <- rbind(test_crop, test)
}


test_crop <- test_crop %>% rename(variable = parameter1)


forest_str_by_crop_code <- statby_crop %>% 
  pivot_wider(values_from = value, names_from = crop_code) %>% 
  inner_join(test_crop) %>% 
  dplyr::select(-expression) 


forest_str_by_crop_code %>%
  write_csv(here::here("data/forest_str_by_crop_code.csv"))


forest_str_by_crop_code %>%  flextable(col_keys = c("variable","CA1","CA2","CA3",
                         "SJ1","SJ2", "statistic", "p.value")) %>% 
  colformat_double(digits = 3)

```

## Exploratory plots by variable

### Composistion (% *Q. pyrenaica*)
```{r}
ggbetweenstats(data = forest_str, x = crop_code,
                              type = "np", pairwise.comparisons = TRUE, 
                              pairwise.display = "s",
                              y = per_qp)
```

### Abundance 
```{r}
ggbetweenstats(data = forest_str, x = crop_code,
                              type = "np", pairwise.comparisons = TRUE, 
                              pairwise.display = "s",
                              y = abundance)
```

### ab_juvenile
```{r}
ggbetweenstats(data = forest_str, x = crop_code,
                              type = "np", pairwise.comparisons = TRUE, 
                              pairwise.display = "s",
                              y = ab_juvenile)
```

### ab_adult
```{r}
ggbetweenstats(data = forest_str, x = crop_code,
                              type = "np", pairwise.comparisons = TRUE, 
                              pairwise.display = "s",
                              y = ab_adult)
```

### tree_height_max
```{r}
ggbetweenstats(data = forest_str, x = crop_code,
                              type = "np", pairwise.comparisons = TRUE, 
                              pairwise.display = "s",
                              y = tree_height_max)
```

### tree_height_mean
```{r}
ggbetweenstats(data = forest_str, x = crop_code,
                              type = "np", pairwise.comparisons = TRUE, 
                              pairwise.display = "s",
                              y = tree_height_mean)
```

### tree_height_median
```{r}
ggbetweenstats(data = forest_str, x = crop_code,
                              type = "np", pairwise.comparisons = TRUE, 
                              pairwise.display = "s",
                              y = tree_height_median)
```

### tree_height_q75
```{r}
ggbetweenstats(data = forest_str, x = crop_code,
                              type = "np", pairwise.comparisons = TRUE, 
                              pairwise.display = "s",
                              y = tree_height_q75)
```

### tree_height_q90
```{r}
ggbetweenstats(data = forest_str, x = crop_code,
                              type = "np", pairwise.comparisons = TRUE, 
                              pairwise.display = "s",
                              y = tree_height_q90)
```

### ba
```{r}
ggbetweenstats(data = forest_str, x = crop_code,
                              type = "np", pairwise.comparisons = TRUE, 
                              pairwise.display = "s",
                              y = ba)
```


